### 2024-09-23 data-visualization.qmd
20240920 - remove extra line
20240923 - faceting

"Showing the data" involves showing the data that matters the most.
"Reducing the clutter" involves removing non-data things that obscure the data—for example, extraneous gridlines, tick marks, data markers (e.g., symbols to distinguish between series), and complex shadings (e.g., textured or filled gradients).
"Integrating the graphics and text" involves using headline titles, clear and useful labels (instead of legends), and helpful annotations.
Headline or newspaper-like titles are titles that are succinct with active phasing and that indicate the take-away message (e.g., "Quarterbacks Threw Fewer Touchdowns in 2024 than in Previous Years").
In terms of labels, @Schwabish2021 advocates to label the data directly instead of using a legend.
In terms of helpful annotations, you can provide additional text that helps explain the data (e.g., peaks or valleys, outliers, or other variations that deserve explanation), including how to interpret the chart.
"Avoiding the spaghetti chart" means avoiding packed charts with too much information that makes them difficult to interpret.
Spaghetti charts are lines with many lines that, make the plot look like a bunch of spaghetti.
However, @Schwabish2021 also advocates against using charts of other types that are complicated and difficult to interpret due to too much information, such as complicated maps or bar plots with too many colors, icons, or bars.
If there are too many lines or series, @Schwabish2021 advocates breaking it up into multiple charts (i.e., facets, trellis charts, or small multiples).
"Starting with gray" refers to the idea of using gray as the default color for most lines/points/bars, so that you can use a color to highlight the lines/points/bars of interest.
In addition, as noted by @Schwabish2021, it is important to treat data as objectively as possible and not to present figures in a biased way as to mislead.

@@

>>>>

"Showing the data" involves showing the data that matters the most.
"Reducing the clutter" involves removing non-data things that obscure the data—for example, extraneous gridlines, tick marks, data markers (e.g., symbols to distinguish between series), and complex shadings (e.g., textured or filled gradients).
"Integrating the graphics and text" involves using headline titles, clear and useful labels (instead of legends), and helpful annotations.
Headline or newspaper-like titles are titles that are succinct with active phasing and that indicate the take-away message (e.g., "Quarterbacks Threw Fewer Touchdowns in 2024 than in Previous Years").
In terms of labels, @Schwabish2021 advocates to label the data directly instead of using a legend.
In terms of helpful annotations, you can provide additional text that helps explain the data (e.g., peaks or valleys, outliers, or other variations that deserve explanation), including how to interpret the chart.
"Avoiding the spaghetti chart" means avoiding packed charts with too much information that makes them difficult to interpret.
Spaghetti charts are lines with many lines that, make the plot look like a bunch of spaghetti.
However, @Schwabish2021 also advocates against using charts of other types that are complicated and difficult to interpret due to too much information, such as complicated maps or bar plots with too many colors, icons, or bars.
If there are too many lines or series, @Schwabish2021 advocates [breaking it up into multiple charts](#sec-faceting) (i.e., facets, trellis charts, or small multiples).
An example of [faceted charts](#sec-faceting) is depicted in @fig-facetedScatterplot.
"Starting with gray" refers to the idea of using gray as the default color for most lines/points/bars, so that you can use a color to highlight the lines/points/bars of interest.
In addition, as noted by @Schwabish2021, it is important to treat data as objectively as possible and not to present figures in a biased way as to mislead.

@@

## Faceting {#sec-faceting}

```{r}
#| label: fig-facetedScatterplot
#| fig-cap: "Faceted Scatterplot."

ggplot2::ggplot(
  data = player_stats_seasonal_offense %>% 
    filter(position_group %in% c("QB","RB","WR","TE")),
  aes(
    x = age,
    y = fantasy_points)) +
  geom_point() +
  geom_smooth() +
  scale_x_continuous(
    expand = c(0,0)
  ) +
  scale_y_continuous(
    expand = c(0,0),
    lim = c(0,NA)
  ) +
  labs(
    x = "Player's Age (years)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Age"
  ) +
  theme_bw() +
  facet_wrap(vars(position_group)) # facet by position_group
```
